---
title: "PS2 Labor Economics [2020]"                      
subtitle: "R Version"
author: "Jonas Veit [jveit@mail.uni-mannheim.de]"
output: 
  html_document:
    toc: true
    toc_depth: 1
    number_sections: false
---
\
\
\

\
\
\
\
![wqewqewqeqwewqewqe text](/home/jonas/Downloads/Uni_Mannheim_Siegel.svg)
\
\
\
\
\

# Estimating the intensive margin labor supply elasticity

\
\
\
Load helpful packages and the cps dataset
```{r setup, include=TRUE, message=FALSE}
library(tidyverse)
library(haven)
library(ggthemes)
library(gridExtra)

cps = read_dta(file = '~/Documents/Programming/R/Labor Economics [2020]//Data for PS2/cps-1984-2010.dta')

```
\
\
\

### $a)$
```{r}
summary(cps)
```

```{r,message=FALSE}
mean(cps$nKids)
mean(cps$hisp)
```
The average number of childen is $0.653$

The share of hispanic women in this data is $15.5$%
\
\
\

### $b)$
```{r , message=FALSE}
hourly_wage = cps$incwage / (52*cps$uhrswork)

mean(hourly_wage,na.rm = TRUE)
sd(hourly_wage,na.rm = TRUE)
```
Since there are NaN's in our hourly wage vector we have to use **na.rm = TRUE**
\
\
$mean(hourly\_wage) = 10.091$
\
\
$sd(houry\_wage) = 13.277$
\
\
There are missing values because the column $uhrswork$ contains zero rows.
\
Dividing by zero produces NaN's. You can calculate the amount of zeros by executing the following command.
```{r}
sum(cps$uhrswork == 0)
```
Compare it with the number of NaN's in the $hourly\_wage$ vector.
```{r}
sum(hourly_wage == 'NaN')
```
\
\
\

### $c)$
```{r}
mean(cps$emp_ind)
mean(cps$emp_ind[cps$nKids == 0])
mean(cps$emp_ind[cps$nKids == 1 & 2])

```
The overall rate of labor force participation is or $74.41$%
\
\
The rate of labor force participation among women who have no children is or $78.59$%
\
\
The rate for 1 or 2 children is or $71.78$%
\
\
The rate of labor force participation among women with 10 children is very high because of the amount of women who have 10 children.
```{r}
cps %>% 
  filter(nKids == 10) %>%
  select(emp_ind)

```
we can clearly see that there are just 10 women with 10 children in this dataset and 6 of them are participating in the labor force i.e $60$%
\
\
\

### $d)$
\

we want to estimate the model $$\textrm{uhrswork}_i = \beta_0 + \beta_1 \textrm{wage}_i + \varepsilon_i$$
```{r}
regression1 = lm(cps$uhrswork ~ hourly_wage);regression1
sum_reg1 = summary(regression1);summary(regression1)[4]
```
The estimate for $\hat{\beta}_1$ is $0.06117$
\
\
The standard error for $\hat{\beta}_1$ is $0.0014$ and the t-statistic is $41.20$
\
We can see that $Pr(>|t|)$ is $0$ that means our p value is 0
\
\
According to the estimated $\hat{\beta}_1$ an one dollar increase in wage should result in an increase of hours worked of 0.0611
\
\
Now estimate the elasticity of hours worked with respect to wage
```{r}
avg_hourly_wage = mean(hourly_wage,na.rm = TRUE)
avg_uhrswork = mean(cps$uhrswork)
regression1$coefficients[2]*(avg_hourly_wage/avg_uhrswork)
```

The estimated elasticity $e$ is equal to $0.0199$
\
\
\

### $e)$
\
Now we want to estimate the log regression $$\log(\textrm{uhrswork}_i) = \beta_0 + \beta_1 \log(\textrm{wage}_i) + \varepsilon_i$$
```{r}
log_uhrswork = log(cps$uhrswork)
log_wage = log(hourly_wage)
log_regression1 = lm(log_uhrswork ~ log_wage);log_regression1
sum_logreg1 = summary(log_regression1);sum_logreg1[4]
```
The estimate for $\hat{\beta}_1$ is $0.08245$
\
\
The standard error for $\hat{\beta}_1$ is $0.00069$ and the t-statistic is $119.2211$
\
We can see that $Pr(>|t|)$ is $0$ that means our p value is 0
\
\
$e_{reg1} = 0.01992475$ \ vs. \ $e_{log(reg1)} = 0.08245$
\
\
If the wage increases by $10$% we would estimate the effect on hours worked to be $10 \times 0.08245 = 0.8245$
\
\
\

### $f)$
\
I am combining $log\_wage$ and $age$ into one dataframe for easier usage.
\
$log\_hours$ contains NaN's that means that we have to use the command: **use = 'complete.obs'**
```{r}
cov_dataframe = data.frame(log_wage,cps$age)
cov(cov_dataframe,use = 'complete.obs')
0.0053*(cov(log_wage,cps$age,use = 'complete.obs')/var(log_wage,na.rm = TRUE))
```
\
$$OVB = \beta_{age} \frac{Cov(log\_wage,age)}{\tilde{S}^2_{log\_wage}} = 0.0053 \frac{2.219}{0.997} = 0.012$$
\
We estimate the OVB to be around 0.012 i.e a elasticity change of 0.012
\
\
\

### $g)$
\
$$\hat{\beta} = \beta + \beta_{age} \frac{Cov(log\_wage,age)}{\tilde{S}^2_{log\_wage}} \Leftrightarrow \beta = \hat{\beta} -  \beta_{age} \frac{Cov(age,log\_wage)}{\tilde{S}^2_{log\_wage}} = 0.08245 - 0.012 = 0.07045$$
\
\
\

### $h)$
\
We want to estimate the model
\
$$\log(\textrm{uhrswork}_i) = \beta_0 + \beta_1 \log(\textrm{wage}_i) +\beta_2\textrm{edu_years} + \beta_3\textrm{age} + \beta_4 \textrm{age}^2 + \beta_5 \textrm{year} + \beta_6 \textrm{hisp} + \varepsilon_i$$
```{r}
age_squared = cps$age^2
log_regression2 = lm(data = cps,formula = log_uhrswork ~ log_wage + edu_yrs + age + age_squared + year + hisp);log_regression2
```
\
The estimated labor supply elasticity does not stay the same. We are now controlling for other variables. Therefore we are able to adress the omitted variable bias.
\
\
\

### $i)$
\
Here we create a new DataFrame with all women with no childen which we access in the $lm()$ command.
\
*** Moreover we create new vectors for $\log(wage)$ , $\log(uhrswork)$ and $\textrm{age}^2$ conditioned on $nKids == 0$
\
Note that I used a different approach to create the DataFrame and vectors using the package 'dplyr' which belongs to the 'tidyverse' package that we loaded at the beginning
\
\
We could have also used the following base-R commands:
\
$$\textrm{cps_nochildren = cps \$ [cps \$ nKids == 0,]}$$
\
\
$$\textrm{log_wage_nochildren} = \log(\textrm{cps_nochildren \$ incwage / (52*cps_nochildren \$ uhrswork}))$$
\
$$...$$
\
\
But using 'dplyr' makes the code much easier to read!
\
```{r}
cps_nochildren = cps %>%
  filter(nKids == 0)

log_wage_nochildren = cps %>%
  filter(nKids == 0) %>%
  select(incwage,uhrswork) %>%
  transmute(log(incwage/(52*uhrswork))) %>%
  pull()


log_uhrswork_nochildren = cps %>%
  filter(nKids == 0) %>%
  select(uhrswork) %>%
  transmute(log(uhrswork)) %>%
  pull()

age_squared_nochildren = cps_nochildren %>%
  select(age) %>%
  transmute(age^2) %>%
  pull()

```


```{r}
log_regression_nochildren = lm(data = cps_nochildren,formula = log_uhrswork_nochildren ~ log_wage_nochildren + edu_yrs + age + age_squared_nochildren + year + hisp);log_regression_nochildren
```
\
```{r}
cps_children = cps %>%
  filter(nKids > 1)

log_wage_children = cps %>%
  filter(nKids > 1) %>%
  select(incwage,uhrswork) %>%
  transmute(log(incwage/(52*uhrswork))) %>%
  pull()


log_uhrswork_children = cps %>%
  filter(nKids > 1) %>%
  select(uhrswork) %>%
  transmute(log(uhrswork)) %>%
  pull()

age_squared_children = cps_children %>%
  select(age) %>%
  transmute(age^2) %>%
  pull()
```

```{r}
log_regression_children = lm(data = cps_children,formula = log_uhrswork_children ~ log_wage_children + edu_yrs + age + age_squared_children + year + hisp);log_regression_children
```
\
\
\

### $j)$
\
```{r}

```
\
\
\
\
\

# Minimum Wage Application
\
\
\
Load the corresponding dataset.
```{r}
dubelesterreich_minwage = read_dta(file = "~/Documents/Programming/R/Labor Economics [2020]/Data for PS2/dubelesterreich_minwage.dta")

dubelesterreich_empdata = read_dta(file = "~/Documents/Programming/R/Labor Economics [2020]/Data for PS2/dubelesterreich_empdata.dta")

dubelesterreich_empdata_contig_minwage = read_dta(file = "~/Documents/Programming/R/Labor Economics [2020]/Data for PS2/dubelesterreich_empdata_contig_minwage.dta")
```
\
\
\

### $1)$ Prepare the dataset

\
Create the meanwage vector and append it to the dataframe.
\
```{r}
minwage = rep(0,length(dubelesterreich_minwage$st_mw))

for (i in 1:length(t(minwage))){
  if (is.na(dubelesterreich_minwage$st_mw[i] == TRUE)){
    minwage[i] = dubelesterreich_minwage$fed_mw[i] 
  } else if (dubelesterreich_minwage$st_mw[i] > dubelesterreich_minwage$fed_mw[i]){
    minwage[i] = dubelesterreich_minwage$st_mw[i]
  } else if (dubelesterreich_minwage$st_mw[i] == dubelesterreich_minwage$fed_mw[i]){
    minwage[i] = dubelesterreich_minwage$st_mw[i]
  } else {
    minwage[i] = dubelesterreich_minwage$fed_mw[i]
  }
}

dubelesterreich_minwage = dubelesterreich_minwage %>%
  add_column(minwage)
```
\
```{r}
summary(dubelesterreich_minwage$st_mw)
summary(dubelesterreich_minwage$fed_mw)
summary(minwage)
```


\
$a)  \  \ \ \textrm{&} \ \ \ b)$
\
$\min = 3.350$
\
$\max = 7.930$
\
\
\
Save the minwage vector as a file.
```{r}
write_dta(as.data.frame(minwage),path = "~/Documents/Programming/R/Labor Economics [2020]/minwage_clean.dta")
```
\
\
$c)$
\
```{r}


mean_minwage = dubelesterreich_minwage %>%
  group_by(year) %>%
  arrange(year) %>%
  summarise(mean = mean(minwage))

mean_fedwage = dubelesterreich_minwage %>%
  group_by(year) %>%
  arrange(year) %>%
  summarise(mean_fed = mean(fed_mw))

mean_minwage_fed = data.frame(mean_fedwage,mean_minwage) %>%
  select(-year.1) %>%
  pivot_longer(cols = c("mean_fed","mean"));colnames(mean_minwage_fed)[3] = "means"

mean_minwage_plot = ggplot(mean_minwage) +
  geom_path(aes(x = year , y = mean),color = "deepskyblue3") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  ggtitle(label = "Average minimum wage by year") +
  ylab("") +
  xlab("Year") +
  theme_stata()

mean_fed_minwage_plot = ggplot(mean_minwage_fed,aes(color = name)) +
  geom_path(aes(x = year , y = means)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  ylab(label ="") +
  xlab(label = "Year") +
  ggtitle(label = "Average minimum wage and federal minimum wage by year") +
  scale_color_manual(name="",labels=c("Average minimum wage","Federal minimum wage"),values=c("deepskyblue3","red")) +
  theme_stata()
```
\
```{r , multi.col=TRUE}
mean_minwage_plot


mean_fed_minwage_plot
```
\
\
\
$d)$
\
Filter for year = 2000
\
```{r}
dubelesterreich_empdata_2000 = dubelesterreich_empdata %>%
  filter(year == 2000)


mean_emp_rest = dubelesterreich_empdata_2000 %>%
  group_by(state) %>%
  arrange(state) %>%
  summarise(mean_emp_rest = mean(emp_rest))

mean_emp_tot = dubelesterreich_empdata_2000 %>%
  group_by(state) %>%
  arrange(state) %>%
  summarise(mean_emp_tot = mean(emp_tot))

mean_emp_totemp = data.frame(mean_emp_rest,mean_emp_tot) %>%
  select(-c("state","state.1"))

mean_emp_totemp_plot = ggplot(mean_emp_totemp) +
  geom_point(aes(x = mean_emp_tot , y = mean_emp_rest),color = "deepskyblue3") +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 5)) +
  theme_stata()

mean_emp_totemp_plot
```
\
\
```{r}
logmean_emp_rest = dubelesterreich_empdata_2000 %>%
  group_by(state) %>%
  arrange(state) %>%
  summarise(logmean_emp_rest = mean(logemp_rest))

logmean_emp_tot = dubelesterreich_empdata_2000 %>%
  group_by(state) %>%
  arrange(state) %>%
  summarise(logmean_emp_tot = mean(logemp_tot))

logmean_emp_totemp = data.frame(logmean_emp_rest,logmean_emp_tot) %>%
  select(-c("state","state.1"))

logmean_emp_totemp_plot = ggplot(logmean_emp_totemp) +
  geom_point(aes(x = logmean_emp_tot , y = logmean_emp_rest),color = "deepskyblue3") +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 5)) +
  theme_stata()

logmean_emp_totemp_plot
```

\
\
\

# Estimating the Effect of Minimum Wages
\
\
Load the corresponding dataset.
\
I have also removed all other variables created for Part 1 of this Problem Set to have a cleaner workspace.
\
```{r}
rm(list = ls())

empdata_minwage = read_dta(file = "~/Documents/Programming/R/Labor Economics [2020]/Data for PS2/dubelesterreich_empdata_contig_minwage.dta")
```
\
\
\

### $a)$
\
\
Equation $(1)$ in the Dube, Lester and Reich article is the following
\
\
$$\ln_{it} = \alpha + \eta\ln(MW_{it}) + \delta\ln(y_{it})^{TOT} + \gamma\ln(pop_{it}) + \phi_i + \tau_t + \varepsilon_{it}$$
\
\
\
First we want to estimate
\
\

$$\log(\textrm{earnings_rest}) = \alpha + \eta\log(\textrm{minwage}) + \delta\log(\textrm{earnings_tot}) + \varepsilon_{it}$$
\
```{r}
log_regression1 = lm(data = empdata_minwage, formula = logearnings_rest ~ logminwage + logearnings_tot);log_regression1

summary(log_regression1)[4]
```
\
$\hat{\eta} = 0.5042$
\
\
The coefficient is statistically significant with a t-value of $94.46$ and a p-value of $0$
\
\
\

### $b)$
\
```{r}
attach(empdata_minwage)

period = empdata_minwage %>%
  select(period)

dummy_dataframe = as.data.frame(matrix(nrow = length(t(period)),ncol = length(min(period):max(period))))
colnames(dummy_dataframe) = paste("period",1:length(dummy_dataframe),sep = "")


for (i in 1:length(min(period):max(period))){
   k = (i + min(period) -1)
   period_dummy = paste("period",i,sep = "")
   dummy_dataframe[,i] = (assign(period_dummy,as.numeric(period == k))) 
}



lm(data = dummy_dataframe,formula = logearnings_rest ~ logminwage + logearnings_tot + .)
```



